---
title: "R Notebook"
output: html_notebook
---

PRELIM DATA CHECK
```{r}
# Load in libraries
library(tidyverse)
library(ggplot2)
library(lubridate)
library(skimr)
library(psych)
library(Hmisc)
library(corrplot)
library(naniar)
library(GGally)
library(janitor)
library(scales)

# Load in clean data sets for journeys, customers
journeys <- read_csv("journeys_clean.csv")
customers <- read_csv("customers_clean.csv")

# Quick check both data frames
head(journeys)
head(customers)

# Check missing values
journeys %>% gg_miss_var()
customers %>% gg_miss_var()

# Summaries of basic descriptives of numeric data
journeys %>% psych::describe()
customers %>% psych::describe()
```

DROP UNNECESSARY COLUMNS
```{r}
# Journeys
journeys %>%
  select(-purchase_own, -purchase_any, -purchase_competitor) -> journeys

# Customers
customers %>%
  select(-...1) -> customers

```


UNIVARIATE - BASIC SUMMARY STATISTICS
```{r}
# Journey duration
journeys %>%
  summarise(mean_duration = mean(duration),
            med_duration = median(duration),
            dev_duration = sd(duration),
            min_duration = min(duration),
            max_duration = max(duration))

# Age
customers %>%
  summarise(mean_age = mean(Age),
            med_age = median(Age),
            dev_age = sd(Age),
            min_age = min(Age),
            max_age = max(Age))

# Children per household
customers %>%
  summarise(mean_children = mean(children_per_household),
            med_children = median(children_per_household),
            dev_children = sd(children_per_household),
            min_children = min(children_per_household),
            max_children = max(children_per_household))

```


UNIVARIATE - COUNT USERS, JOURNEYS, CONVERSIONS
```{r}
# Users - 9678 unique 
journeys %>%
  select(UserID) %>%
  n_distinct()

# Purchase journeys - 29012 unique
journeys %>%
  select(journeyID) %>%
  n_distinct()

# Bookings at own agency - 162
journeys %>%
  filter(purchase == "Own") %>%
  select(UserID) %>%
  n_distinct()

# Bookings at competitor agencies - 2830
journeys %>%
  filter(purchase == "Competitor") %>%
  select(UserID) %>%
  n_distinct()

```

UNIVARIATE - CUSTOMER DEMOGRAPHICS
```{r}
# Region - West and South most common
customers %>%
  tabyl(region) %>%
  drop_na() %>%
  arrange(n) %>%
  ggplot() + geom_col(aes(y = n, x = reorder(region, n)), fill = "#1a6b89") + coord_flip() +
  theme_classic(base_size = 16) +
  labs(
    title = "Distribution of Customer Regions",
    x = "Region",
    y = "Customers")
  
# Municipality size - Most common between 20,000 ~ 100,000 (< 500,000)
customers %>%
  tabyl(municipality_size) %>%
  drop_na() %>%
  arrange(n) %>%
  ggplot() + geom_col(aes(y = n, x = reorder(municipality_size, n)), fill = "#1a6b89") + coord_flip() +
  theme_classic(base_size = 16) +
  labs(
    title = "Distribution of Municipality Sizes",
    x = "Size Category",
    y = "Customers")

# Household size - Over 60% of all customers are either single or have 2 people in household
# Additional ~14% have 3 people
customers %>% 
  tabyl(household_size) %>%
  adorn_pct_formatting(digits = 2) %>% 
  adorn_totals()

# Gender - 60% female
customers %>% 
  tabyl(GenderID) %>%
  adorn_pct_formatting(digits = 1) %>% 
  adorn_totals()

# Employment - Majority of customers are employed (regular salaried) or retired
customers %>%
  tabyl(employment) %>%
  drop_na() %>%
  arrange(n) %>%
  ggplot() + geom_col(aes(y = n, x = reorder(employment, n)), fill = "#48c78a") + coord_flip() +
  theme_classic(base_size = 16) +
  labs(
    title = "Distribution of Employment Categories",
    x = "Employment Category",
    y = "Customers")

# Income - 16% unreported; 37% of remaining users earned average to 1-2 times average wage
customers %>% 
  tabyl(income) %>%
  adorn_pct_formatting(digits = 1) %>% 
  adorn_totals()

customers %>%
  tabyl(income) %>%
  #drop_na() %>%
  arrange(n) %>%
  ggplot() + geom_col(aes(y = n, x = reorder(income, n)), fill = "#48c78a") + coord_flip() +
  theme_classic(base_size = 16) +
  labs(
    title = "Distribution of Income",
    x = "Income Category",
    y = "Customers")


# Social class - 47% of customers are middle to upper middle class
customers %>% 
  tabyl(social_class) %>%
  adorn_pct_formatting(digits = 1) %>% 
  adorn_totals()

customers %>%
  tabyl(social_class) %>%
  #drop_na() %>%
  arrange(n) %>%
  ggplot() + geom_col(aes(y = n, x = reorder(social_class, n)), fill = "#48c78a") + coord_flip() +
  theme_classic(base_size = 16) +
  labs(
    title = "Distribution of Social Class",
    x = "Social Class Category",
    y = "Customers")

# Education - Notable amount customers graduated MBO 2, 3, 4 of MBO before 1998
customers %>% 
  tabyl(education) %>%
  adorn_pct_formatting(digits = 1) %>% 
  adorn_totals()

customers %>%
  tabyl(education) %>%
  #drop_na() %>%
  arrange(n) %>%
  ggplot() + geom_col(aes(y = n, x = reorder(education, n)), fill = "#1a6b89") + coord_flip() +
  theme_classic(base_size = 16) +
  labs(
    title = "Distribution of Education Levels",
    x = "Education Category",
    y = "Customers")

# Life stage - 30% empty nesters, 14% mature singles, 14% mature families
customers %>%
  tabyl(life_stage) %>%
  drop_na() %>%
  arrange(n) %>%
  ggplot() + geom_col(aes(y = n, x = reorder(life_stage, n)), fill = "#1a6b89") + coord_flip() +
  theme_classic(base_size = 16) +
  labs(
    title = "Distribution of Life Stages",
    x = "Life Stage Category",
    y = "Customers")

customers %>% 
  tabyl(life_stage) %>%
  adorn_pct_formatting(digits = 1) %>% 
  adorn_totals()


```

LOAD IN ADDITIONAL DATA SETS
```{r}
purchase_by_device <- read_csv("purchase_by_device.csv")
purchase_touchpoints <- read_csv("purchase_touchpoints.csv")
own_purchases <- read_csv("own_purchases.csv")
comp_purchases <- read_csv("comp_purchases.csv")
journey_durations <- read_csv("journey_durations.csv")

```


BIVARIATE - JOURNEYS AND CONVERSIONS BY DEVICE, TOUCHPOINT CATEGORIES (CIC)
```{r}
# Journeys and conversions classified by device type, touchpoint category
plot_data <- journeys %>%
  count(device_type, touch_point_category, purchase) %>% 
  group_by(touch_point_category, device_type) %>% 
  mutate(percent = n/sum(n))

ggplot(plot_data, aes(x = device_type, y = percent, fill = purchase)) + 
  geom_col(position = "fill") + 
  geom_label(aes(label = percent(percent)), position = "fill", color = "white", vjust = 1, show.legend = FALSE) +
  scale_y_continuous(labels = percent) +
  facet_grid(~touch_point_category) +
  scale_fill_manual(values=c("#1a6b89", "#dc6f6f", "#48c78a")) +
  labs(
    x = "Device Type",
    y = "Amount of Touchpoints") + theme_classic(base_size = 20)
  
```


BIVARIATE - JOURNEYS AND CONVERSIONS BY SPECIFIC TOUCHPOINT TYPES
```{r}
# Bar chart - touchpoint category type distributions
journeys %>%
  tabyl(touchpoint_group) %>%
  drop_na() %>%
  arrange(n) %>%
  ggplot() + geom_col(aes(y = n, x = reorder(touchpoint_group, n)), fill = "#dc6f6f") + coord_flip() +
  theme_classic(base_size = 16) +
  labs(
    title = "Distribution of Customer-Initiated Touchpoint Types",
    x = "Platform",
    y = "Journeys")

# Journeys and conversions classified by device type, touchpoint type (coded)
plot_data_2 <- journeys %>%
  drop_na() %>%
  count(touch_point_channel, touchpoint_group, purchase) %>% 
  group_by(touchpoint_group, touch_point_channel) %>% 
  mutate(percent = n/sum(n))

ggplot(plot_data_2, aes(x = touch_point_channel, y = percent, fill = purchase)) + 
  geom_col(position = "fill") + 
  geom_label(aes(label = percent(percent)), position = "fill", color = "white", vjust = 1, show.legend = FALSE) +
  scale_y_continuous(labels = percent) +
  facet_grid(~touchpoint_group) +
  scale_fill_manual(values=c("#1a6b89", "#dc6f6f", "#48c78a")) +
  theme_classic(base_size = 18) +
  labs(
    x = "Channel",
    y = "Amount of Touchpoints")
```


```{r}
journeys %>%
  filter(purchase == "Own") %>%
  ggplot() + geom_bar(aes(x = purchase_own, fill = touchpoint_group)) +
  theme_classic() +
  scale_fill_manual(values=c("#dc6f6f", "#f78a8a", "#F5BDBD", "#1a6b89", "#48c78a")) +
  labs(
      title = "No. Touchpoints Led to Purchase",
      x = "Own Purchases",
      y = "Amount of Touchpoints")
  


ggplot(journeys, aes(x = purchase_own, fill = touchpoint_group)) +
    geom_bar() +
    theme_classic() +
    scale_fill_manual(values=c("#dc6f6f", "#f78a8a", "#F5BDBD", "#1a6b89", "#48c78a")) +
    labs(
      title = "No. Touchpoints Led to Purchase",
    x = "Purchases",
    y = "Amount of Touchpoints")


table(journeys$touchpoint_group)




```

BIVARIATE - CLASSIFYING BY JOURNEY DURATION, TOUCHPOINT TYPE
```{r}
# Histograms - journey durations grouped by touchpoint type (for top used touchpoint platforms)
journeys %>%
  filter(purchase == "Own" | purchase == "Competitor") %>%
  filter(touchpoint_group == "Accommodations" | touchpoint_group == "Travel - Competitor") %>%
  drop_na() %>%
  ggplot() + geom_histogram(aes(duration), fill = "#dc6f6f") +
  facet_wrap(touchpoint_group~., scales="free") +
  theme_classic(base_size = 20) +
  labs(
    x = "Seconds",
    y = "Journeys")
  
```


COMBINE JOURNEYS AND CUSTOMERS DATA SETS
```{r}
# Join both data sets on user ID
journeys %>%
  left_join(customers, by = c("UserID" = "UserID")) %>%
  {.} -> journeys.customers
```

JOURNEY AND DEMOGRAPHIC ATTRIBUTES FOR CONVERTED CUSTOMERS
```{r}
# Demographic distributions for competitor bookings
journeys.customers %>%
  filter(purchase == "Competitor") %>%
  drop_na() %>% 
  #filter(touchpoint_group == "Flights" | touchpoint_group == "Information") %>%
  ggplot() + geom_histogram(aes(employment), fill = "#1a6b89", bins = 30) +
  facet_wrap(touchpoint_group~., scales="free") +
  theme_classic(base_size = 20) +
  labs(
    x = "Employment Category",
    y = "Journeys")

# Demographic distributions for own bookings
journeys.customers %>%
  filter(purchase == "Own") %>%
  drop_na() %>% 
  #filter(touchpoint_group == "Flights" | touchpoint_group == "Information") %>%
  ggplot() + geom_histogram(aes(employment), fill = "#48c78a", bins = 30) +
  facet_wrap(touchpoint_group~., scales="free") +
  theme_classic(base_size = 20) +
  labs(
    x = "Employment Category",
    y = "Journeys")
```


```{r}
table(journeys$touch_point_category)


journeys %>%
  drop_na() %>%
  filter(purchase == "Own") %>%
  filter(touchpoint_group == "Accommodations") %>%
  count()

```


